Supporting Information

### 2015-12-15 si.Rmd
Stefan and Helene's comments from lab meeting Added acknowledgements (fixes #32) and more biology to the conclusion (fixes #31) Shorter title (fixes #25)

---
title: "Supporting Information"
csl: csl/the-plant-journal.csl
date: "\\vspace{0 mm}"
header-includes:
  \usepackage{graphicx}
  \usepackage{caption}
  \usepackage{fixltx2e}
  \usepackage{rotating}
  \renewcommand{\thefigure}{S\arabic{figure}}
  \renewcommand{\thetable}{S\arabic{table}}
documentclass: tufte-handout
classoption: a4paper
output:
  rmarkdown::tufte_handout:
    dev: CairoPDF
    keep_tex: yes
    fig_crop: no
bibliography: bib/papersLib.bibtex
---

@@

\clearpage
\begin{figure*}
  \includegraphics{data/isFigure.pdf}
\end{figure*}
\captionof{figure}{RNA \emph{in situ} hybridization analysis to confirm the specific expression patterns of four genes detected in the RNA sequencing dataset. Expression of \emph{LOC\_Os09g27730} (a--d), \emph{LOC\_Os01g04670} (e--h), \emph{LOC\_Os10g04270} (i--l) and \emph{LOC\_Os10g05990} (m--p) was analysed at the rachis meristem (a, e, i and m), primary branch meristem (b, f, j and n), elongating primary branch meristem plus axillary meristem (c, g, k and o) and spikelet meristem (d, h, l and p) stages. Each gene was detected in the same stage by \emph{in situ} hybridization as by RNA sequencing.}

@@

```{r sf_madsTree, fig.height = 9.535, fig.cap = paste0("Differential expression of genes encoding putative MIKC proteins in \\emph{A. thaliana}, \\emph{Solanum lycopersicum} and \\emph{Oryza sativa} ssp. \\emph{japonica}. The tree is a neighbour-joining tree estimated from genetic distance of amino acid sequences (see Methods S1). Named proteins starting with AT\\_ or OS\\_ are from Arabidopsis or rice respectively, with synonyms in parentheses, and other proteins are labelled with their unique identifiers.", "\\newline ", "The colour of the label represents the Log\\textsubscript{2}-fold change (L\\textsubscript{2}FC) in expression of the gene encoding that protein between undifferentiated and differentiated meristems. Where there is no coloured label, the adjusted \\emph{p}-value associated with the difference in expression was greater than 0.1, or the gene for the protein in question was excluded from the differential expression analysis by \\texttt{DESeq2}, which may occur because of low read counts, an extreme count outlier or the gene failing independent filtering \\citep{@Love:2014do}.", "\\newline ", "For rice, L\\textsubscript{2}FC was estimated from the RM and SM samples reported here. For \\emph{A. thaliana}, L\\textsubscript{2}FC was estimated using the inflorescence meristem and floret meristem samples reported by \\citep{Mantegazza:2014di}. The sympodial inflorescence meristem and floret meristem samples generated by \\citep{Park:2012dg} were used to estimate L\\textsubscript{2}FC values for tomato genes. Clades are labelled following the nomenclature presented by \\citep{Arora:2007fa}.")}
print(sf_madsTree)
```

@@

# Supporting experimental procedures

## Methods S1: Comparative analysis of MADS family

The primary sequences of *Arabidopsis thaliana*, *Solanum lycopersicum* and *Oryza sativa* MADS proteins were downloaded by querying Phytozome 10.3 [@Goodstein:2012ho] for the PANTHER family PTHR11945 ('MADS BOX PROTEIN') via the `biomaRt` interface [@Thomas:2003db; @Durinck:2009ki]. An initial alignment of all proteins was performed with `Clustal Omega` using a full distance matrix [@Sievers:2011fn]. To clean the alignment, poorly conserved regions where greater than `r minpcnongap` percent of sequences had gaps or less than `r minpcident` percent of pairwise comparisons between sequences had identical residues were removed. To recover MADS genes related to the Arabidopsis MIKC clade, a UGPMA tree was estimated from the cleaned alignment and the tree was cut at the lowest height that returned a clade including all type II Arabidopsis MADS proteins listed in the Arabidopsis transcription factor database on the Arabidopsis Gene Regulatory Information Server [@Yilmaz:2011ew], with as few type I proteins as possible. Protein sequences in this clade that were longer than `r minProtLength` amino acids were realigned using `Clustal Omega`, and the alignment was cleaned as described above and used to estimate a neighbour-joining tree with the `bionj` algorithm implemented in the `ape` package [@Paradis:2004dv] rooted to type I MADS sequences (`r og`). L~2~FC change expression values for tomato and Arabidopsis genes were estimated from samples that have previously been described [@Mantegazza:2014di; @Park:2012dg] using the same method described above for the rice samples. The tree was plotted using the `ggtree` extension for `ggplot2` [https://github.com/GuangchuangYu/ggtree; @Wickham:2009tn].

>>>>

---
title: "Supporting Information"
csl: csl/the-plant-journal.csl
date: "\\vspace{0 mm}"
header-includes:
  \usepackage{graphicx}
  \usepackage{caption}
  \usepackage{fixltx2e}
  \usepackage{rotating}
  \usepackage{siunitx}
  \renewcommand{\thefigure}{S\arabic{figure}}
  \renewcommand{\thetable}{S\arabic{table}}
documentclass: tufte-handout
classoption: a4paper
output:
  rmarkdown::tufte_handout:
    dev: CairoPDF
    keep_tex: yes
    fig_crop: no
bibliography: bib/papersLib.bibtex
---

@@

\clearpage
\begin{figure*}
  \includegraphics{data/isFigure.pdf}
\end{figure*}
\captionof{figure}{RNA \emph{in situ} hybridization analysis to confirm the specific expression patterns of four genes detected in the RNA sequencing dataset. Expression of \emph{LOC\_Os09g27730} (a--d), \emph{LOC\_Os01g04670} (e--h), \emph{LOC\_Os10g04270} (i--l) and \emph{LOC\_Os10g05990} (m--p) was analysed at the rachis meristem (a, e, i and m), primary branch meristem (b, f, j and n), elongating primary branch meristem plus axillary meristem (c, g, k and o) and spikelet meristem (d, h, l and p) stages. Each gene was detected in the same stage by \emph{in situ} hybridization as by RNA sequencing. The scale bars represent \SI{100}{\micro\metre}.}

@@

```{r sf_madsTree, fig.height = 9.535, fig.cap = paste0("Differential expression of genes encoding putative MIKC proteins in \\emph{A. thaliana}, \\emph{Solanum lycopersicum} and \\emph{Oryza sativa} ssp. \\emph{japonica}. The tree is a neighbour-joining tree estimated from genetic distance of amino acid sequences (see Methods S1). Named proteins starting with AT\\_ or OS\\_ are from Arabidopsis or rice respectively, with synonyms in parentheses, and other proteins are labelled with their unique identifiers.", "\\newline ", "The colour of the label represents the Log\\textsubscript{2}-fold change (L\\textsubscript{2}FC) in expression of the gene encoding that protein between undifferentiated and differentiated meristems. Where there is no coloured label, the adjusted \\emph{p}-value associated with the difference in expression was greater than 0.1, or the gene for the protein in question was excluded from the differential expression analysis by \\texttt{DESeq2}, which may occur because of low read counts, an extreme count outlier or the gene failing independent filtering (Love \\emph{et al}., 2014).", "\\newline ", "For rice, L\\textsubscript{2}FC was estimated from the RM and SM samples reported here. For \\emph{A. thaliana}, L\\textsubscript{2}FC was estimated using the inflorescence meristem and floret meristem samples reported by Mantegazza \\emph{et al}. (2014). The sympodial inflorescence meristem and floret meristem samples generated by Park \\emph{et al}. (2012) were used to estimate L\\textsubscript{2}FC values for tomato genes.")}
print(sf_madsTree)
```

@@

# Methods S1

## Comparative analysis of the MADS family

The primary sequences of *Arabidopsis thaliana*, *Solanum lycopersicum* and *Oryza sativa* MADS proteins were downloaded by querying Phytozome 10.3 [@Goodstein:2012ho] for the PANTHER family PTHR11945 ('MADS BOX PROTEIN') via the `biomaRt` interface [@Thomas:2003db; @Durinck:2009ki]. An initial alignment of all proteins was performed with `Clustal Omega` using a full distance matrix [@Sievers:2011fn]. To clean the alignment, poorly conserved regions where greater than `r minpcnongap` percent of sequences had gaps or less than `r minpcident` percent of pairwise comparisons between sequences had identical residues were removed. To recover MADS genes related to the Arabidopsis MIKC clade, a UGPMA tree was estimated from the cleaned alignment and the tree was cut at the lowest height that returned a clade including all type II Arabidopsis MADS proteins listed in the Arabidopsis transcription factor database on the Arabidopsis Gene Regulatory Information Server [@Yilmaz:2011ew], with as few type I proteins as possible. Protein sequences in this clade that were longer than `r minProtLength` amino acids were realigned using `Clustal Omega`, and the alignment was cleaned as described above and used to estimate a neighbour-joining tree with the `bionj` algorithm implemented in the `ape` package [@Paradis:2004dv] rooted to type I MADS sequences (`r og`). L~2~FC change expression values for tomato and Arabidopsis genes were estimated from samples that have previously been described [@Mantegazza:2014di; @Park:2012dg] using the same `STAR`–`DESeq2` pipeline described in the main text for the rice samples [@Dobin:2012fg; @Love:2014do]. The tree was plotted using the `ggtree` extension for `ggplot2` [https://github.com/GuangchuangYu/ggtree; @Wickham:2009tn], with clade labels following the nomenclature presented by [@Arora:2007fa].
